---
title: "How Wealth Affects Olympic Performance"
author: "Annabelle Haryanto, Ethan Hsu, Kassy Chaput"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

**P02: Exploratory Data Analysis**

```{r, echo = FALSE}
# Summary Info Calculations
source("../source/summary_info.R")
options(scipen = 999)
```

*Dynamic paragraph:*

We sought to find and display a correlation between a countries’ Gross Domestic Product (GDP) and their Olympic success which we measured through total medal count. We have observed `r summary_info$summary_num_countries` different countries throughout this process. We have found that the average GDP from 2021 across all countries is approximately `r round(summary_info$summary_avg_gdp)` and the average medal count for each country as of 2021 is approximately `r round(summary_info$summary_avg_medal_count)` . We have further observed that approximately `r summary_info$summary_above_avg_gdp` countries have a GDP higher than the world average and that approximately `r summary_info$summary_above_avg_medals` countries have an above average number of total medals won at the Olympics.

*Table:*
```{r, echo = FALSE}
# Aggregate Table
source("../source/AggregateTable.R")
```


*Observations:*

From this table we learn that the average medal count across all countries is `r average_medals` and that the average national gross domestic product (GDP) is approximately `r round(median_2020_GDP)`. These values are crucial to our analysis because we previously calculated how many countries are above these averages, so including these values provides the threshold in which we are considering. It also provides context to our data visualizations because it can be easily observed whether a certain country in our visualizations is above or below this threshold. Throughout our analysis we discovered the United States to be well above average in both of these areas so we have also included that `r (round(usa_gold_percentage, digits = -2)) * 100` percent of total gold medals that have been won by the United States.

```{r, echo = FALSE}
# GDP Map

source("../source/gdp map.R")
ggplotly(gdp_map)
```

*Observations:*

This map is included to visualize the Gross Domestic Product (GDP) per capita by each country from data collected at the end of 2021. This can be used to achieve a broad overview of the economy in every country across the world. The countries with the darkest green color have the highest GDP such as the United States, Australia, and a few countries in Europe. Those with a lighter green color have a lower GDP production such as many of the countries in both Africa and South America. The countries with high GDP (darker color) can also further be compared to the bar chart below to help determine if a correlation between high GDP and Olympic success exists.  

```{r, echo = FALSE}
# Bar Chart

source("../source/Bar_Chart.R")
ggplotly(top_performers)
```

*Observations:*

This chart is to visualize the top 10 performing countries (in medal count) over the course of Olympic history. This chart can be used to quickly analyze locations around the world that perform best at the Olympics. The graphic reveals that the United States has the most medals of any country at 5219. We can also see that they have about 2 times as many medals as the next best Olympic team. Another observation we can make from the plot is that Europe seems to be the highest performing continent in Olympic history. South America, Africa, and Asia are left out of the graphic entirely, signaling to us that these continents have no Olympic powerhouses. 

```{r, echo = FALSE}
# United States Time Graph

source("../source/medals time frame.R")
ggplotly(medals_US)
```

*Observations:*

As seen from the previous two data visualizations, the United States is among the top in both GDP production and Olympic success, this time graph goes further in depth tracking the United States' Olympic success over time. This reveals that the United States has had relatively stable success in the Olympics until 1980. After 1980, the United States has had varied success, some years total medal count is nearly 750, where as others it is closer to 250. This difference may be due to the Olympics being held every 2 years and staggering Winter and Summer Olympics after 1986. We can see that in the year 1992, the United States won the most medals and after that, the graph fluctuates over the years until 2016.

