---
title: "How Wealth Affects Olympic Performance"
author: "Annabelle Haryanto, Ethan Hsu, Kassy Chaput"
date: "11/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

**P02: Exploratory Data Analysis**

*Dynamic paragraph:*

*Table:*
```{r, echo = FALSE}
# Aggregate Table

#source("../data/AggregateTable.R")
#average_gold
```

*Three Charts:*

```{r, echo = FALSE}
# GDP Map

source("../source/gdp map.R")
ggplotly(gdp_map)
```

This map is included to visualize the Gross Domestic Product (GDP) per capita by each country from data collected at the end of 2021. This can be used to achieve a broad overview of the economy in every country across the world. The countries with the darkest green color have the highest GDP such as the United States, Australia, and a few countries in Europe. Those with a lighter green color have a lower GDP production such as many of the countries in both Africa and South America. The countries with high GDP (darker color) can also further be compared to the bar chart below to help determine if a correlation between high GDP and Olympic success exists.  

```{r, echo = FALSE}
# Bar Chart

source("../source/Bar_Chart.R")
ggplotly(top_performers)
```

This chart is to visualize the top 10 performing countries (in medal count) over the course of Olympic history. This chart can be used to quickly analyze locations around the world that perform best at the Olympics. The graphic reveals that the United States has the most medals of any country at 5219. We can also see that they have about 2 times as many medals as the next best Olympic team. Another observation we can make from the plot is that Europe seems to be the highest performing continent in Olympic history. South America, Africa, and Asia are left out of the graphic entirely, signaling to us that these continents have no Olympic powerhouses. 

```{r, echo = FALSE}
# United States Time Graph

source("../source/medals time frame.R")
ggplotly(medals_US)
```


